Centralizers of higher enveloping algebras and their large $N$ limits

### 2019-05-07 frame.tex
Greg's edit

\def\brian{\textcolor{blue}{BW: }\textcolor{blue}}
\def\owen{\textcolor{magenta}{OG: }\textcolor{magenta}}
\def\mahmoud{\textcolor{green}{MZ: }\textcolor{green}}

@@

The centralizer ${\frak Z}_{E_n}(f)$ is the $E_n$ algebra $\clies(\fg, B)$.
\end{thm}

By using Knudsen's results, we have the following consequence.

@@

When $k$ is $\RR$ or $\CC$, 
it is modeled by the locally constant factorization algebra that assigns
\[
\clies(\fg \otimes \Omega^*(U), \cliels( \fg \otimes \Omega^*_c(U)))
\]
to each open set $U \subset \RR^n$.
\end{cor}

This result generalizes the fact that $\Hoch^*(U\fg,U\fg) \simeq \clies(\fg, U\fg)$, since the associative algebra $U\fg$ is modeled by the locally constant one-dimensional factorization algebra that assigns $\cliels (\fg \otimes \Omega^*_c(U))$ to each open $U \subset \RR$.

Having an explicit presentation for the centralizer of a map of $E_n$ algebras allows for immediate calculations of factorization homology.
Indeed, for $M$ a closed $n$-dimensional manifold,
\[
\int_M {\frak Z}_{E_n}(\UU_n \rho) \simeq \clies(\fg \otimes \Omega^*(M), \cliels( \fh \otimes \Omega^*(M))).
\]
When $M = S^n$, we thus see
\[
\int_{S^n} {\frak Z}_{E_n}(\UU_n \rho) \simeq \clies(\fg[\epsilon_n], \cliels( \fh[\epsilon_n]))
\]
as $\Omega^*(S^n) \simeq \CC[\epsilon_n]$ where $\epsilon_n$ is a formal variable with $|\epsilon_n| = n$ and $\epsilon_n^2 = 0$.
In the case that $n=1$, this recovers a theorem of Tamarkin-Tsygan \cite{TT} on the Hochschild cohomology of differential operators in the case of a group $G$ (see Example \ref{eg: TT}).

@@

The explicit factorization algebras provide convenient methods for computing factorization homology.
As shown in \owen{exact reference} in \cite{Knudsen}, 
the enveloping $E_n$ algebras are homotopy $O(n)$-fixed points and hence determine factorization algebras on unoriented $n$-dimensional manifolds.

\section{Enveloping $E_n$ algebras of Lie algebras}

@@

An alternative characterization is found in Theorem 5.3.1.30 of~\cite{LurieHA}.

@@

We record a key fact drawn from \cite{LurieHA} and~\cite{FrancisHH}, which generalizes this situation.

@@

See Remark 7.3.5.3 of \cite{LurieHA} and Section 2 of \cite{FrancisHH}.
Concretely, this claim arises from the fact that as a module over $A$ as an $E_n$ algebra, 
$M \in \Mod^{E_n}_A(\cC)$ is equipped with a commuting family of actions by $A$ parametrized by $\RR^n \setminus \{0\}$.
This action factors through the left action of the $E_1$ algebra $\int_{S^{n-1} \times \RR} A$ arising from factorization homology over the complement of the origin,
which integrates all those actions and through which they all factor.
As shown in Theorem 7.5.3.1 of \cite{LurieHA} and earlier in Proposition 3.16 of \cite{FrancisHH}, 
this algebra $\int_{S^{n-1} \times \RR} A$ is equivalent to ${\rm Free}(1_\cC)$, 
the image of the unit object $1_\cC \in \cC$ under the left adjoint ${\rm Free}$ to the forgetful functor $\Mod^{E_n}_A(\cC) \to \cC$.

@@

Let $\fg$ be a dg Lie algebra, let $B$ be an $E_n$ algebra, and let $f: \UU_n \fg \to B$ be a map of $E_n$ algebras.
Then the results of the preceding section imply
\begin{align*}
{\frak Z}_{E_n}(f) &\simeq {{\rm Mor}}_{\Mod^{E_n}_{\UU_n}}(\UU_n \fg, B_f) \\
&\simeq {\rm Mor}_{\LMod_{\int_{S^{n-1}} \UU_n}}(\UU_n \fg, B_f)
\end{align*}
where 
\begin{itemize}
\item $B_f$ denotes the object in $\Mod^{E_n}_{\UU_n}$ or, equivalently, in $\LMod_{\int_{S^{n-1}}\UU_n}$ determined by~$f$,
\item ${\rm Mor}_{\Mod^{E_n}_{\UU_n}}(\UU_n \fg, B)$ denotes the classifying object for maps from $\UU_n$ to $B_f$ in the $\infty$-category $\Mod^{E_n}_{\UU_n}$ (i.e., an internal mapping object),
\item and ${\rm Mor}_{\LMod_{\int_{S^{n-1}} \UU_n}} (\UU_n \fg, B_f)$ denotes the classifying object for maps from ${\UU_n}$ to $B$ in the $\infty$-category $\LMod_{\int_{S^{n-1}}\UU_n}$.
\end{itemize}
Observe now that $\int_{S^{n-1}}\UU_n$ is an $E_1$ algebra equivalent to the dg algebra $U(\fg \otimes k[\epsilon_{n-1}])$.
This is because the rational $(n-1)$-sphere is formal, and hence its cochains are quasi-isomorphic to its cohomology $k[\epsilon_{n-1}]$.
Using the canonical inclusion $i: U\fg \to U(\fg \otimes k[\epsilon_{n-1}])$,
we find
\[
k \otimes_{U\fg} U(\fg \otimes k[\epsilon_{n-1}]) \simeq \Sym(\fg[1-n]).
\]
Recall now that $\UU_n \fg \simeq \Sym(\fg[1-n])$ as $\fg$-modules.
Base change thus implies that we have an equivalence
\[
{\rm Mor}_{\LMod_{\int_{S^{n-1}} \UU_n}} (\UU_n \fg, B_f)  \simeq {\rm Mor}_{U\fg} (k, B_f)
\]
of underlying chain complexes.
(That is, we have an equivalence after applying forgetful functors.)
The right hand side is precisely Lie algebra cochains of $\fg$ with coefficients in~$B_f$, as desired.

\brian{do you like this example?}
\begin{eg}\label{eg: TT}
Consider the case $n=1$ and $B = U \fg$ where $f$ is simply the identity map. 
The Hochschild cohomology of $U \fg$ can be expressed in terms of Lie algebra cohomology as $\Hoch^*(U\fg,U\fg) \simeq \clies(\fg, U\fg)$. 
On the other hand, our model for the derived center ${\frak Z}(U(\fg))$ of $U \fg$ is the one-dimensional factorization algebra assigning 
\[
\clies(\fg \otimes \Omega^*(U), \cliels( \fg \otimes \Omega^*_c(U)))
\]
to each open set $U \subset \RR$.
In particular, the factorization homology of the derived center along the closed manifold $S^1$ can be computed as
\[
\int_{S^1} {\frak Z}(U(\fg)) \simeq \clies(\fg[\epsilon], \cliels( \fg[\epsilon])),
\]
where $\epsilon$ is a formal variable of degree~$1$. 
\brian{We should compare with Tamarkin-Tsygan's noncommutative claculus (cf. section 2.6).
We could exhibit Theorem 2.7.1 explicitly in this case for the $E_n$ case with enveloping algebras.}
\end{eg}

@@

\begin{thm}
For $A$ a unital dg algebra over a characteristic zero field $k$,
there is a quasi-isomorphism
\beqn\label{lqt}
\lambda_A: \cliels(\cL(A)) \xto{\simeq} \Sym(\cC(A))
\eeqn
of dg cocommutative Hopf algebras.
\end{thm}

@@

\begin{cor}
There is a natural transformation
\[
\lambda: \cliels(\cL(-)) \Longrightarrow \Sym(\cC(-))
\]
between these two functors from the dg category of unital dg algebras to the dg category of dg cocommutative Hopf algebras over~$k$.
It is a quasi-isomorphism on every algebra.
\end{cor}

For our purposes, a variant is more useful, but this variant follows by easy, formal arguments.

@@

We explain here how to construct a factorization algebra using cyclic homology,
much as Lie algebra homology leads to an explicit factorization Lodayalgebra modeling the enveloping $E_n$ algebra. 
Throughout this section, let $k$ denote the field $\RR$ or $\CC$; 
by mimicking \cite{Knudsen}, one could also produce an explicit model over~$\QQ$.

\begin{dfn}
Let $A$ be a dg algebra over $k$, and let $\cA = A \otimes \Omega^*_c$ denote the associated cosheaf of dg algebras on~$\RR^n$.
Let $\cC_A$ denote the precosheaf $\Cyc_*(\cA)[1]$ of cochain complexes on $\RR^n$,
which assigns to each open set $U$, the 1-shifted cyclic chains on $\cA(U)$.
\end{dfn}

Note that for two algebras $B$ and $B'$,
\[
\Cyc_*(B) \oplus \Cyc_*(B') \simeq \Cyc_*(B \times B') .
\] 
The corresponding statement for ordinary Hochschild homology is induced by the projections $B \leftarrow B \times B' \to B'$, and any map of algebras that induces a quasi-isomorphism on Hochschild homology induces one on cyclic homology as well.
(See, for instance, \S2.2.12 of \cite{LodayCyclic}.)
Hence, for any finite collection of disjoint opens $U_1$, \dots, $U_k$, 
we have
\[
\cC_A(U_1 \sqcup \cdots \sqcup U_k) \simeq \cC_A(U_1) \oplus \cdots \oplus \cC_A(U_k).
\]
Moreover, for any nested sequence of inclusions $V_1 \subset V_2 \subset \cdots$ with $V = \bigcup_i V_i$,
there is a quasi-isomorphism
\[
{\rm hocolim}(\cC_A(V_1) \to \cC_A(V_2) \to \cdots) \simeq \cC_A(V).
\]
(See E.2.1.1 of \cite{LodayCyclic}.)
In this sense $\cC_A$ behaves like a cosheaf, up to quasi-isomorphism. 
This property underlies the following claim.

>>>>

\def\brian{\textcolor{blue}{BW: }\textcolor{blue}}
\def\owen{\textcolor{magenta}{OG: }\textcolor{magenta}}
\def\mahmoud{\textcolor{green}{MZ: }\textcolor{green}}
\def\greg{\textcolor{red}{GG: }\textcolor{red}}

@@

The centralizer ${\frak Z}_{E_n}(f)$ is the $E_n$ algebra $\clies(\fg, B)$.
\end{thm}
\greg{shall explain/remind somewhere what is the $E_n$-structure there: namely induced by $E_n$-structure of $B$ and com mutliplication in $\clies$, namely convolution between an $E_\infty$-coalgebra and an $E_n$-algebra}
By using Knudsen's results, we have the following consequence.

@@

When $k$ is $\RR$ or $\CC$, 
it is modeled by the locally constant factorization algebra that assigns
\[
\clies(\fg \otimes \Omega^*(U), \cliels( \fg \otimes \Omega^*_c(U)))
\]
to each open set $U \subset \RR^n$.
\end{cor}
\greg{shall relate this to $P_n$-center of $\fg$. that shall also (but I have to think to the proof more), at least up to formality gives the $P_{n+1}$-structure on it. }
This result generalizes the fact that $\Hoch^*(U\fg,U\fg) \simeq \clies(\fg, U\fg)$, since the associative algebra $U\fg$ is modeled by the locally constant one-dimensional factorization algebra that assigns $\cliels (\fg \otimes \Omega^*_c(U))$ to each open $U \subset \RR$.

Having an explicit presentation for the centralizer of a map of $E_n$ algebras allows for immediate calculations of factorization homology.

Indeed, as shown  in \cite{Knudsen}, 
the enveloping $E_n$ algebras are homotopy $O(n)$-fixed points (in other words $\mathrm{Disk}_n$-algebras in the sense of~\cite{AF}) and hence determine factorization algebras on unoriented $n$-dimensional manifolds.

Therefore, for $M$ a closed $n$-dimensional manifold, one has 
\[
\int_M {\frak Z}_{E_n}(\UU_n \rho) \simeq \clies(\fg \otimes \Omega^*(M), \cliels( \fh \otimes \Omega^*(M))).
\]
When $M = S^n$, we thus see
\[
\int_{S^n} {\frak Z}_{E_n}(\UU_n \rho) \simeq \clies(\fg[\epsilon_n], \cliels( \fh[\epsilon_n]))
\]
as $\Omega^*(S^n) \simeq \CC[\epsilon_n]$ where $\epsilon_n$ is a formal variable with $|\epsilon_n| = n$ and $\epsilon_n^2 = 0$.
In the case that $n=1$, this recovers a theorem of Tamarkin-Tsygan \cite{TT} on the Hochschild cohomology of differential operators in the case of a group $G$ (see Example \ref{eg: TT}).

@@

Again, the explicit factorization algebras provide convenient methods for computing factorization homology.

\greg{need to add a notation/convention somewhere including that $\Sigma^n$ will be shift functor, convention for grading and so on}
\section{Enveloping $E_n$ algebras of Lie algebras}

@@

An alternative characterization is found in Theorem 5.3.1.30 of~\cite{LurieHA} or Proposition 6.22 in~\cite{GTZ3}.

@@

We record a key fact drawn from \cite{LurieHA},~\cite{FrancisHH}  and \cite{GG-Notes}, which generalizes this situation.

@@

See Remark 7.3.5.3 of \cite{LurieHA}, Coroallary 13 in~\cite{GG-Notes} and Section 2 of \cite{FrancisHH}.
Concretely, this claim arises from the fact that as a module over $A$ as an $E_n$ algebra, 
$M \in \Mod^{E_n}_A(\cC)$ is equipped with a commuting family of actions by $A$ parametrized by $\RR^n \setminus \{0\}$.
This action factors through the left action of the $E_1$ algebra $\int_{S^{n-1} \times \RR} A$ arising from factorization homology over the complement of the origin,
which integrates all those actions and through which they all factor.
As shown in Theorem 7.5.3.1 of \cite{LurieHA} and earlier in Proposition 3.16 of \cite{FrancisHH}, 
this algebra $\int_{S^{n-1} \times \RR} A$ is equivalent to ${\rm Free}(1_\cC)$, 
the image of the unit object $1_\cC \in \cC$ under the left adjoint ${\rm Free}$ to the forgetful functor $\Mod^{E_n}_A(\cC) \to \cC$.

@@

Let $\fg$ be a dg Lie algebra, let $B$ be an $E_n$ algebra, and let $f: \UU_n \fg \to B$ be a map of $E_n$ algebras.
Then the results of the preceding section imply
\begin{align*}
{\frak Z}_{E_n}(f) &\simeq {{\rm Mor}}_{\Mod^{E_n}_{\UU_n}}(\UU_n \fg, B_f) \\
&\simeq {\rm Mor}_{\LMod_{\int_{S^{n-1}} \UU_n}}(\UU_n \fg, B_f)
\end{align*}
where 
\begin{itemize}
\item $B_f$ denotes the object in $\Mod^{E_n}_{\UU_n}$ or, equivalently, in $\LMod_{\int_{S^{n-1}}\UU_n}$ determined by~$f$,
\item ${\rm Mor}_{\Mod^{E_n}_{\UU_n}}(\UU_n \fg, B)$ denotes the classifying object for maps from $\UU_n$ to $B_f$ in the $\infty$-category $\Mod^{E_n}_{\UU_n}$ (i.e., an internal mapping object),
\item and ${\rm Mor}_{\LMod_{\int_{S^{n-1}} \UU_n}} (\UU_n \fg, B_f)$ denotes the classifying object for maps from ${\UU_n}$ to $B$ in the $\infty$-category $\LMod_{\int_{S^{n-1}}\UU_n}$.
\end{itemize}
Observe now that $\int_{S^{n-1}}\UU_n$ is an $E_1$ algebra equivalent to the dg algebra $U(\fg \otimes k[\epsilon_{n-1}])$.
This is because the rational $(n-1)$-sphere is formal, and hence its cochains are quasi-isomorphic to its cohomology $k[\epsilon_{n-1}]$.
Using the canonical inclusion $i: U\fg \to U(\fg \otimes k[\epsilon_{n-1}])$,
we find
\[
k \otimes_{U\fg} U(\fg \otimes k[\epsilon_{n-1}]) \simeq \Sym(\fg[1-n]).
\]
Recall now that $\UU_n \fg \simeq \Sym(\fg[1-n])$ as $\fg$-modules \greg{shall we also need that it is true as $E_n$-algebra one we see right hand side as a $P_n$-algebra in the natrual way and use formality ? I feel we need to say something more to recover the $E_n$-structure below-something like the $E_n$-structure on Lie algebra cochains realizes teh moduel map structure after all equivalences we use}.
Base change thus implies that we have an equivalence
\[
{\rm Mor}_{\LMod_{\int_{S^{n-1}} \UU_n}} (\UU_n \fg, B_f)  \simeq {\rm Mor}_{U\fg} (k, B_f)
\]
of underlying chain complexes.
(That is, we have an equivalence after applying forgetful functors.)
The right hand side is precisely Lie algebra cochains of $\fg$ with coefficients in~$B_f$, as desired.

\brian{do you like this example?} \greg{I do !}
\begin{eg}\label{eg: TT}
Consider the case $n=1$ and $B = U \fg$ where $f$ is simply the identity map. 
The Hochschild cohomology of $U \fg$ can be expressed in terms of Lie algebra cohomology as $\Hoch^*(U\fg,U\fg) \simeq \clies(\fg, U\fg)$. 
On the other hand, our model for the derived center ${\frak Z}(U(\fg))$ of $U \fg$ is the one-dimensional factorization algebra assigning 
\[
\clies(\fg \otimes \Omega^*(U), \cliels( \fg \otimes \Omega^*_c(U)))
\]
to each open set $U \subset \RR$.
In particular, the factorization homology of the derived center along the closed manifold $S^1$ can be computed as
\[
\int_{S^1} {\frak Z}(U(\fg)) \simeq \clies(\fg[\epsilon], \cliels( \fg[\epsilon])),
\]
where $\epsilon$ is a formal variable of degree~$1$. 
\brian{We should compare with Tamarkin-Tsygan's noncommutative claculus (cf. section 2.6).
We could exhibit Theorem 2.7.1 explicitly in this case for the $E_n$ case with enveloping algebras.}
\end{eg}

@@

\begin{thm}[\cite{LQ}]
For $A$ a unital dg algebra over a characteristic zero field $k$,
there is a quasi-isomorphism
\beqn\label{lqt}
\lambda_A: \cliels(\cL(A)) \xto{\simeq} \Sym(\cC(A))
\eeqn
of dg cocommutative Hopf algebras.
\end{thm}

@@

\begin{cor}
There is a natural transformation
\[
\lambda: \cliels(\cL(-)) \Longrightarrow \Sym(\cC(-))
\]
between these two functors from the dg category of unital dg algebras to the dg category of dg cocommutative Hopf algebras over~$k$.
It is a quasi-isomorphism on every algebra.
\end{cor}
\greg{why not state (in a remark) this corollary and the next one as syaing this is an equivalence in the $\infty$-categories of functor ?}
For our purposes, a variant is more useful, but this variant follows by easy, formal arguments.

@@

We explain here how to construct a factorization algebra using cyclic homology,
much as Lie algebra homology leads to an explicit factorization algebra modeling the enveloping $E_n$ algebra. 
Throughout this section, let $k$ denote the field $\RR$ or $\CC$; 
by mimicking \cite{Knudsen}, one could also produce an explicit model over~$\QQ$.

\begin{dfn}
Let $A$ be a dg algebra over $k$, and let $\cA = A \otimes \Omega^*_c$ denote the associated cosheaf of dg algebras on~$\RR^n$.
Let $\cC_A$ denote the precosheaf $\Cyc_*(\cA)[1]$ of cochain complexes on $\RR^n$,
which assigns to each open set $U$, the 1-shifted cyclic chains on $\cA(U)$.
\end{dfn}
\greg{Here we shall either recall our model for cyclic chains, or first define the $\infty$-categorical universal statement and then precise a nice model. By the universal model, I mean define cyclic as the homotopy $S^1$-fixed point for the action on Hochschild chains. This will make the construction more transparent ?}
Note that for two algebras $B$ and $B'$,
\[
\Cyc_*(B) \oplus \Cyc_*(B') \simeq \Cyc_*(B \times B') .
\] 
The corresponding statement for ordinary Hochschild homology is induced by the projections $B \leftarrow B \times B' \to B'$, and any map of algebras that induces a quasi-isomorphism on Hochschild homology induces one on cyclic homology as well.
(See, for instance, \S2.2.12 of \cite{LodayCyclic}.)
Hence, for any finite collection of disjoint opens $U_1$, \dots, $U_k$, 
we have
\[
\cC_A(U_1 \sqcup \cdots \sqcup U_k) \simeq \cC_A(U_1) \oplus \cdots \oplus \cC_A(U_k).
\]
Moreover, for any nested sequence of inclusions $V_1 \subset V_2 \subset \cdots$ with $V = \bigcup_i V_i$,
there is a quasi-isomorphism
\[
{\rm hocolim}(\cC_A(V_1) \to \cC_A(V_2) \to \cdots) \simeq \cC_A(V).
\]
(See E.2.1.1 of \cite{LodayCyclic}.)
In this sense $\cC_A$ behaves like a cosheaf, up to quasi-isomorphism. 
This property underlies the following claim.
